---
title: "Feature tour"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Feature tour}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(printtree)
```

This article walks through the output features of `printtree`.

```{r demo-tree}
demo <- file.path(tempdir(), "printtree-feature-tour")
if (dir.exists(demo)) unlink(demo, recursive = TRUE, force = TRUE)

dir.create(file.path(demo, "R"), recursive = TRUE)
dir.create(file.path(demo, "data", "raw"), recursive = TRUE)
dir.create(file.path(demo, "logs"), recursive = TRUE)
dir.create(file.path(demo, "empty"), recursive = TRUE)

file.create(file.path(demo, "R", "helpers.R"))
file.create(file.path(demo, "README.md"))
file.create(file.path(demo, "data", "raw", "sales.csv"))
file.create(file.path(demo, "logs", "debug.log"))
file.create(file.path(demo, "test_cache"))
```

## Count summaries

By default, `print_rtree()` ends with a displayed directory/file count.

```{r count-summary}
print_rtree(demo, max_depth = 2)
```

Set `count_footer = FALSE` for compact output.

```{r compact-output}
print_rtree(demo, max_depth = 1, count_footer = FALSE)
```

## Pattern ignores

The `ignore` argument still supports exact basenames, but with
`ignore_type = "auto"` it also treats wildcard entries as glob patterns.

```{r glob-ignore}
print_rtree(demo, ignore = c("*.log", "test_*"), max_depth = 2)
```

You can opt into regular expression matching for more control.

```{r regex-ignore}
print_rtree(demo, ignore = "^(README|test_)", ignore_type = "regex", max_depth = 1)
```

## Prune empty directories

Use `prune = TRUE` to hide directories with no displayable children after
ignores and depth limits are applied.

```{r prune}
print_rtree(demo, ignore = "*.log", prune = TRUE)
```

## Quiet capture

Use `quiet = TRUE` with `return_lines = TRUE` when you want to work with the
tree programmatically.

```{r quiet-capture}
lines <- print_rtree(demo, return_lines = TRUE, quiet = TRUE)
head(lines, 4)
```

## Text and Markdown export

`write_tree()` writes the same tree output to a text or Markdown file. Parent
directories are created automatically by default.

```{r write-tree}
tree_md <- file.path(tempdir(), "printtree-feature-tour-output", "tree.md")
write_tree(demo, tree_md, format = "md", title = "Feature Tour Tree")
readLines(tree_md, n = 8)
```

## Mermaid and Graphviz diagrams

`tree_to_mermaid()` and `tree_to_dot()` convert the tree to diagram text that
renders in Quarto documents: Mermaid flowcharts render natively in `{mermaid}`
chunks (and on GitHub), while DOT graphs render in `{dot}` chunks when
Graphviz is installed. Both accept the same filtering options as the printed
tree (`ignore`, `max_depth`, `show_hidden`, `prune`).

```{r mermaid-preview, echo = FALSE, results = "asis"}
diagram <- tree_to_mermaid(demo, max_depth = 2)
diagram <- gsub("&", "&amp;", diagram, fixed = TRUE)
diagram <- gsub("<", "&lt;", diagram, fixed = TRUE)
diagram <- gsub(">", "&gt;", diagram, fixed = TRUE)
cat('<pre class="mermaid">\n', diagram, '\n</pre>\n', sep = "")
cat(paste0(
  '<script type="module">\n',
  "import mermaid from 'https://cdn.jsdelivr.net/npm/mermaid@12/dist/mermaid.esm.min.mjs';\n",
  "mermaid.initialize({startOnLoad:true,securityLevel:'strict',theme:'default'});\n",
  '</script>\n'
))
```

Use `view_mermaid()` for the same rendered preview from an interactive R
session. The viewer includes pan/zoom controls for large trees and browser-side
SVG, PNG, and JPEG download buttons. The raw Mermaid source remains available
from `tree_to_mermaid()`.

```{r mermaid-source, eval = FALSE}
view_mermaid(demo, pan_zoom = TRUE)
cat(tree_to_mermaid(demo, max_depth = 2))
```

For automation, `save` writes a clean rendered image without the viewer toolbar.
PNG, JPEG, and PDF export require the suggested `webshot2` package and a
Chrome/Chromium browser. A missing extension is appended automatically; an
explicit extension must match `save`.

```{r mermaid-save, eval = FALSE}
view_mermaid(demo, file = "tree.png", save = "png", open = FALSE)
view_mermaid(demo, file = "tree.pdf", save = "pdf", open = FALSE)
```

These examples are not evaluated while building the vignette, so package and
CRAN checks never launch a browser. Mermaid itself is loaded from a public CDN,
which means the viewer and rendered export require an internet connection.

```{r dot}
cat(tree_to_dot(demo, max_depth = 2))
```

`write_tree()` can write either diagram format directly, or produce a minimal
Quarto document with the tree embedded as a Mermaid diagram:

```{r write-diagram}
tree_qmd <- file.path(tempdir(), "printtree-feature-tour-output", "tree.qmd")
write_tree(demo, tree_qmd, format = "qmd", title = "Feature Tour Tree",
           max_depth = 2)
readLines(tree_qmd, n = 12)
```

### Diagram extras

The diagram functions have a few more tricks. `git_colors = TRUE` tints nodes
by Git status (amber = modified, green = untracked, blue = staged),
`subgraph = TRUE` wraps each directory in a Mermaid subgraph container, and
`repo_url` makes every node clickable, linking to the file or directory on
GitHub:

```{r mermaid-extras}
cat(tree_to_mermaid(demo, max_depth = 2, git_colors = TRUE, subgraph = TRUE,
                    repo_url = "https://github.com/PrigasG/printtree"))
```

`tree_to_mindmap()` renders the same tree as a radial Mermaid mindmap, and
`tree_to_html()` produces a self-contained HTML page with click-to-expand
directories (no external dependencies):

```{r mindmap}
cat(tree_to_mindmap(demo, max_depth = 2))
```

```{r html, eval = FALSE}
tree_html <- file.path(tempdir(), "printtree-feature-tour-output", "tree.html")
write_tree(demo, tree_html, format = "html", title = "Feature Tour Tree",
           max_depth = 2)
```

Finally, `tree_diff_mermaid()` and `tree_diff_dot()` compare two directory
trees: nodes only in the second tree are green, nodes only in the first are
red and dashed.

```{r diff-demo, eval = FALSE}
cat(tree_diff_mermaid("project-v1", "project-v2"))
```

To use the output in your own Quarto document, generate the diagram once and
paste it into a `{mermaid}` chunk — or let `write_tree()` produce the whole
Quarto document for you:

```{r quarto-pattern, eval = FALSE}
# Generate once, then paste the output into a {mermaid} chunk
# in your Quarto document:
cat(tree_to_mermaid("myproject", max_depth = 2))

# Or write a ready-to-render Quarto document directly:
write_tree("myproject", "tree.qmd", format = "qmd", title = "Project tree")
```

## Git status annotations

When the target folder is inside a Git work tree, `git = TRUE` annotates changed
paths with simple status markers and includes a legend.

```{r git-status, eval = FALSE}
repo <- file.path(tempdir(), "printtree-feature-tour-git")
if (dir.exists(repo)) unlink(repo, recursive = TRUE, force = TRUE)
dir.create(repo, recursive = TRUE)

system2("git", c("-C", repo, "init"), stdout = FALSE, stderr = FALSE)
system2("git", c("-C", repo, "config", "user.email", "test@example.com"))
system2("git", c("-C", repo, "config", "user.name", "Test User"))

file.create(file.path(repo, "tracked.txt"))
system2("git", c("-C", repo, "add", "tracked.txt"), stdout = FALSE, stderr = FALSE)
system2("git", c("-C", repo, "commit", "-m", "initial"), stdout = FALSE, stderr = FALSE)

writeLines("changed", file.path(repo, "tracked.txt"))
file.create(file.path(repo, "new.txt"))

print_rtree(repo, git = TRUE)
```

```{r git-status-fallback, eval = FALSE}
print_rtree("path/to/repo", git = TRUE)
```
